Comparison of Catalogs

### 2024-10-19 compare/compare.qmd
minor fixes

In this script we will compare 2 catalogs, HECATE (@kovlakasHeraklionExtragalacticCatalogue2021) and LCV ([@karachentsevUPDATEDNEARBYGALAXY2013, @karachentsevSTARFORMATIONPROPERTIES2013a])

@@

inner = dt
outer = QTable(ascii.read("../tables/outer_join.ecsv"), masked=True)
hec_not_lcv = QTable(ascii.read("../tables/HEC_not_LCV_join.ecsv"), masked=True)
lcv_not_hec = QTable(ascii.read("../tables/LCV_not_HEC_join.ecsv"), masked=True)

lcv = QTable(ascii.read("../tables/final_table.ecsv"), masked=True)
hec = QTable(ascii.read("../tables/HECATE_LCV.ecsv"), masked=True)
```

The dataset we are going to use for the comparison (inner join) consists of `{python} len(dt)` galaxies and `{python} len(dt.colnames)` columns.

@@

## Completeness of the LCV Catalog

@@

1.  Radial velocities of $$
     V_{LG}<600 \text{ km}\cdot\text{m}^{-1}
    $$ {#eq-lcv-vel}

2.  Distances of $$
       D<11\ \text{Mpc}
       $$ {#eq-lcv-dis}

@@

1.  Completeness within a 10 Mpc radius is difficult to assess due to: - Variability in galaxy properties (luminosity, size, surface brightness, gas content) - Errors in distance measurements (Tully–Fisher method errors of \~20–25%), especially at the 10 Mpc boundary.
    -   Accurate distances are mainly known within \~5 Mpc.
    -   Non-Hubble motions (\~300 km/s)[^1] may make up half of the adopted velocity constraint (@eq-lcv-vel)
    -   Solution for our usage: only keep the galaxies inside a radius of $D=11 Mpc$
    -   HI sources in surveys with low angular resolution: "The presence around our Galaxy of hundreds of high-velocity clouds with low line-of-sight velocities and small W50 widths also provokes the inclusion of false “nearby” dwarf galaxies in the LV"(@karachentsevUPDATEDNEARBYGALAXY2013)
2.  Astro-Spam and Survey Errors: - Automatic surveys produce false detections (e.g., stars misclassified as galaxies, high-velocity clouds mistaken for dwarfs).
3.  Conditional Completeness Estimate: - Galaxies brighter than $M_B^c = -11^m$ or with linear diameters larger than $A_{26} = 1.0$ kpc show (40–60)% completeness. - "among the members of the Local Group (D \< 1 Mpc), only half of the galaxies have absolute magnitudes brighter than −11m. Consequently, more than half of the ultra-faint dwarf companions around normal galaxies, like the Sombrero galaxy (D = 9.3 Mpc), still remain outside our field of view."
4.  Undetected Ultra-Faint Dwarfs: - Many faint dwarf galaxies remain undetected beyond the Local Group. Estimated population of undetected dwarfs could be as large as $10^3–10^4$ within the LV.
5.  Surface Brightness Distribution: - Surface brightness remains consistent across distances, except for ultra-faint dwarfs ($SB <31 \text{mag}\cdot\text{arcsec}^{-2}$). - Faintest dwarf galaxies are detectable only nearby due to their resolution into individual stars.
6.  Luminosity-Size Relationship: - Observations align with cosmological models predicting $L\sim A^3$ (@navarroStructureColdDark1996), though deviations occur for extremely low surface brightness galaxies. - "The deviation from it at the extremely low surface brightness end is due to a systematic overestimation of dwarf galaxy sizes, the brightness profiles of which lie entirely below the Holmberg isophote."

@@

@@

```{python}
#| output: asis
# Creating a dataframe with the lengths of each table
data = {
    "Table": ["Inner join", "Outer join", "LCV", "HECATE", "Unique galaxies in LCV", "Unique Galaxies in Hecate"],
    "Number of galaxies": [len(dt), len(outer), len(lcv), len(hec), len(lcv_not_hec), len(hec_not_lcv)]
}

@@

Completeness (LCV)= `{python} round(len(dt)/len(lcv) *100)` %

@@

Completeness (HECATE)= `{python} round(len(hec_not_lcv)/len(hec)*100)` %

Completeness (LCV)= `{python} round(len(lcv_not_hec)/len(lcv)*100)` %

Combined Completeness =$\frac{\text{Total galaxies in Outer}}{\text{Unique galaxies in HECATE + LCV}}$= `{python} round(len(outer)/(len(lcv_not_hec)+(len(hec_not_lcv))) * 100)` %

@@

```{python}
def table_completeness(dataset_choice, column_name="Dis", description = "Distance", keep_nan = False):
    """
    Creates a table of counts and percentages for each bin based on the chosen dataset (LCV or HEC).

    Parameters:
    - dataset_choice: A string indicating which dataset to use ("lcv" or "hec")
    - column_name: The column name in the datasets to analyze (default is "Dis")
    - description: what to show in the title and the axis (default "Distance")
    """

    if dataset_choice == "lcv":
        main_df, unique_df, label = lcv, lcv_not_hec, "LCV"
    elif dataset_choice == "hec" and hec is not None and hec_not_lcv is not None:
        main_df, unique_df, label = hec, hec_not_lcv, "HEC"
    else:
        raise ValueError("Invalid dataset choice. Please choose 'lcv' or 'hec' with valid datasets.")

@@

    for i in range(len(bin_centers)):
        if main_counts[i] > 0:
            unique_pct = (unique_counts[i] / main_counts[i]) * 100
            inner_pct = (inner_counts[i] / main_counts[i]) * 100
            table_data.append({
                "Bin Start": round(bin_edges[i]),
                "Bin End": round(bin_edges[i+1]),
                f"Unique %": f'{unique_pct:.0f}%',
                "Inner %": f'{inner_pct:.0f}%'
            })
    table_df = pd.DataFrame(table_data)
    return table_df
```

```{python}
#| fig-cap: "Histograms showing the Distance Completeness of the Catalogs"
#| layout-ncol: 2
#| fig-subcap: 
#| - "HECATE"
#| - "LCV"
#| label: fig-dis-comp

@@

main = hec["D"].data
unique = hec_not_lcv["D"].data
counts, bin_edges = np.histogram(main)

@@

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)
ax2.set(xlabel = "Distance [{}]".format(lcv["Dis"].unit), ylabel="Completeness %")
plt.show()

#LCV
main = lcv["Dis"].data
unique = lcv_not_hec["Dis"].data
counts, bin_edges = np.histogram(main)

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "LCV", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LCV", ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Distance for LCV")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100 
###
sns.lineplot(x = bin_edges[1:], y = unique_pct, marker = "o", ax = ax2)

@@

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)
ax2.set(xlabel = "Distance [{}]".format(lcv["Dis"].unit), ylabel="Completeness %")
plt.show()

@@

```{python}
#| fig-cap: "Histograms showing the Type Completeness of the Catalogs"
#| layout-ncol: 2
#| fig-subcap: 
#| - "HECATE"
#| - "LCV"
#| label: fig-type-comp
main = hec["T"].data
main = main[~np.isnan(main)]
unique = hec_not_lcv["T"].data
unique = unique[~np.isnan(unique)]
counts, bin_edges = np.histogram(main, bins = 12)

@@

ax2.set(xlabel = "Types of galaxies", ylabel="Completeness %")
plt.show()
######
#LCV
main = lcv["TType"].data
main = main[~np.isnan(main)]
unique = lcv_not_hec["TType"].data
unique = unique[~np.isnan(unique)]
counts, bin_edges = np.histogram(main)

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "LCV", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LCV", ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Types for LCV")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100

@@

As we can see from the histograms (@fig-dis-comp) and (@fig-type-comp) the sample of unique galaxies of each catalog, gets smaller by an almost constant proportion (Inner join).

This means there is no bias in the selection of the galaxies.

# How are we going to compare the data?

@@

For the distrabution comparisons we will use histograms and Kernel Density Estimate (KDE) plots. The KDE plots visually represent the distribution of data, providing insights into its shape, central tendency, and spread.

@@

$$\text{Percentage change} = \frac{V_{Hecate} - V_{LCV}}{V_{Hecate}}\cdot 100 \%$$

```{python}
def relative_diff(hec, lcv):
    """
    Calculate the relative difference between two sets of values.

    Parameters:
    hec (numpy.ndarray): An array of values representing the HEC dataset.
    lcv (numpy.ndarray): An array of values representing the LCV dataset.

    Returns:
    numpy.ndarray: An array of relative differences between the HEC and LCV datasets.
    The relative difference is calculated as ((HEC - LCV) * 100) / |HEC|.
    Any infinite values in the result are replaced with NaN.
    """
    diff = (hec - lcv) * 100 / np.abs(hec)
    diff[np.isinf(diff)] = np.nan
    return diff
```

@@

| LCV | HECATE | Description |              Pearson Correlation \[-1,1\]              |
|:---------------:|:---------------:|:---------------:|:--------------------:|
| Dis |   D    |  Distance   | `{python} round(compare_data(dt["Dis"],dt["D"])[3],3)` |

@@

plt.close()
fig, ax = plt.subplots()
sns.regplot(x = dt["Dis"].data, y = dt["D"].data,
                 scatter_kws=dict(alpha=0.5, color='blue', edgecolors='white'),
                 line_kws=dict(alpha=0.7, color='red', linewidth=3))
ax.errorbar(dt["Dis"].data, dt["D"].data, yerr = dt["E_D"].data, fmt='none', capsize=2, zorder=1, color = "blue", alpha = 0.3)
results = scipy.stats.linregress(x = dt["Dis"], y=dt["D"])
plt.xlabel("$D_{LCV}$"+" [{}]".format(dt["Dis"].unit))
plt.ylabel("$D_{HEC}$"+" [{}]".format(dt["D"].unit))
plt.legend(["Data",
            "$D_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot D_{LCV}+$"+
                f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.tight_layout()
plt.show()
```

@@

```{python}
rvel_corr = round(compare_data(dt["RVel"],dt["V"])[3],3)
```

|     LCV     |    HECATE    |              Description               |  Linear Correlation  |
|:----------------:|:-----------------:|:----------------:|:----------------:|
| RVel (km/s) |   V (km/s)   |      Heliocentric radial velocity      | `{python} rvel_corr` |
| VLG (km/s)  |              |            Radial velocity             |                      |
|  cz (km/s)  |              |         Heliocentric velocity          |                      |
|             | V_VIR (km/s) | Virgo-infall corrected radial velocity |                      |

@@

ax.errorbar(dt["RVel"].data, dt["V"].data, yerr = dt["E_V"].data, fmt='none', capsize=2, zorder=1, color = "blue")
results = scipy.stats.linregress(x = dt["RVel"], y=dt["V"])
plt.xlabel("$V_{LCV}$"+" [{}]".format(dt["RVel"].unit))
plt.ylabel("$V_{HEC}$"+" [{}]".format(dt["V"].unit))
plt.legend(["Data",
            "$V_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot V_{LCV}+$"+
            f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()
```

@@

```{python}
#| label: fig-vel-pairplot
#| fig-cap: "The correlation Matrix of the Velocities. The lower left triangle is is composed of the scatter plots of the various velocities, the diagonal shows their distrubution and the upper right triangle shows the correlations of the Velocities" 
pairplot_with_correlation(vel_data)
```

\[?\] The close correlation between all of the velocities, could be due to the fact that all of them measure the velocity of each galaxy, but from a different frame of reference.

The strong correlation between the velocities reported by the HECATE and LCV catalogs can be attributed to the fact that both measure the same intrinsic galaxy velocities but reference them to different frames. For instance, velocities can be corrected for the movement of the Earth (heliocentric frame) or the for the Virgo-infall. These corrections shift the measured velocity slightly, but the fundamental measurement remains consistent. This explains why the velocities are highly correlated, as they represent the same physical phenomenon but adjusted for different reference points.

{{<pagebreak>}}

@@

```{python}
#| echo: false

ttype_corr = round(compare_data(dt["TType"], dt["T"])[3], 4)
inc_corr = round(compare_data(dt["inc"], dt["INCL"], sigma = True)[3], 3)
a26_corr = round(compare_data(dt["a26_1"], dt["R1"], sigma =False)[3], 3)
```

|      LCV      |        HECATE        |                        Description                        | Pearson Correlation \[-1,1\] |
|:----------------:|:----------------:|:-----------------:|:----------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |    `{python} ttype_corr`     |
|      inc      |         INCL         |                     Inclination (deg)                     |     `{python} inc_corr`      |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |     `{python} a26_corr`      |

@@

```{python}
#| label: fig-types-compare
#| layout-ncol: 2
#| layout-align: center
#| fig-cap: "Comparison of the Types of galaxies"
#| fig-subcap: 
#| - "All the galaxies, without the correction" 
#| - "Comparison with the correction $|{T_{HECATE}-T_{LCV}}|<7$"

@@

ax.errorbar(dt["TType"].data, dt["T"].data, yerr = dt["E_T"].data, fmt='none', capsize=2, zorder=1, color = "blue")
temp = dt[["TType", "T"]].to_pandas().dropna()
results = scipy.stats.linregress(x = temp["TType"],
                                y = temp["T"])
plt.xlabel("$Type_{LCV}$")
plt.ylabel("$Type_{HEC}$")
plt.legend(["Data",
            "$T_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot T_{LCV}+$"+
            f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()

@@

ax.errorbar(types["TType"].data, types["T"].data, yerr = types["E_T"].data, fmt='none', capsize=2, zorder=1, color = "blue")
results = scipy.stats.linregress(x= types["TType"].data,
                                y = types["T"].data)
plt.xlabel("$Type_{LCV}$")
plt.ylabel("$Type_{HEC}$")
plt.legend(["Data",
            "$T_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot T_{LCV}+$"+
            f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()
```

-   we can assume that the galaxies from the upper left and lower right regions of the plot are classified differently because of this
-   this might explain the large errors E_T of HECATE
-   why we see a difference in the distribution of (@fig-type-comp)

We can remove the "problematic" galaxies by only keeping the ones with: 
$$
|{T_{HECATE}-T_{LCV}}|<\sigma_{T_{LCV}} = 4.8
$$

-   The average uncertainty of the morphological type in the HECATE catalog is $\overline{\delta T} = \pm 1.4$, so the intercept is included in the error.
-   So we can assume that the Distances are the same

@@

sns.histplot(temp["INCL"], kde = True, label = "HECATE")
sns.histplot(temp["inc"], kde = True, label = "LCV")
plt.legend()
plt.xlabel("Inclination of galaxies [deg]")
plt.show()

@@

```{python}
temp.describe()
```

We can see that for values in the range $[\sim 30^\circ,\sim 80^\circ]$, the values of the LCV inclination are higher (@fig-incl-compare). However, since their means, median, min and maxes are similar and the percentage change is practically 0% (mean, median, $\sigma$ = 0 with a range $[-3\%,1\%]$), we can ignore the differences and assume they are the same values.

@@

plt.xlabel("(Major angular diameter$)_{LCV}$")
plt.ylabel("(Semi major axis$)_{HEC}$")
plt.legend(["Data",
            "$R1_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot a_{26,LCV}$"+
            f"${results.intercept:+.2f}$"+
            f"\n $R^2 = {results.rvalue**2*100:.2f}\%$"])
plt.show()
dt["log_a26"] = np.log10(dt["a26_1"].data)
dt["log_R1"] = np.log10(dt["R1"].data)
plt.close()
fig, ax = plt.subplots()
p = sns.regplot(x = dt["log_a26"].data, y = dt["log_R1"].data,
                 scatter_kws=dict(alpha=0.5, color='blue', edgecolors='white'),
                 line_kws=dict(alpha=0.7, color='red', linewidth=3))

>>>>

In this script we will compare 2 catalogs, HECATE (@kovlakasHeraklionExtragalacticCatalogue2021) and LVG ([@karachentsevUPDATEDNEARBYGALAXY2013, @karachentsevSTARFORMATIONPROPERTIES2013a])

@@

inner = dt
outer = QTable(ascii.read("../tables/outer_join.ecsv"), masked=True)
hec_not_lvg = QTable(ascii.read("../tables/HEC_not_LVG_join.ecsv"), masked=True)
lvg_not_hec = QTable(ascii.read("../tables/LVG_not_HEC_join.ecsv"), masked=True)

lvg = QTable(ascii.read("../tables/final_table.ecsv"), masked=True)
hec = QTable(ascii.read("../tables/HECATE_LCV.ecsv"), masked=True)
```

The dataset we are going to use for the comparison (inner join) consists of `{python} len(dt)` galaxies and `{python} len(dt.colnames)` columns with their physical  characteristics.

@@

## Completeness of the LVG Catalog

@@

1.  Radial velocities of

$$
     V_{LG}<600 \text{ km}\cdot\text{m}^{-1}
$$ {#eq-lvg-vel}

1.  Distances of

$$
       D<11\ \text{Mpc}
$$ {#eq-lvg-dis}

@@

1.  Completeness within a 10 Mpc radius is difficult to assess due to: - Variability in galaxy properties (luminosity, size, surface brightness, gas content) - Errors in distance measurements (Tully–Fisher method errors of \~20–25%), especially at the 10 Mpc boundary.
    -   Accurate distances are mainly known within \~5 Mpc.
    -   Non-Hubble motions (\~300 km/s)[^1] may make up half of the adopted velocity constraint (@eq-lvg-vel)
        -   Solution for our usage: only keep the galaxies inside a radius of $D=11 Mpc$
    -   HI sources in surveys with low angular resolution: "The presence around our Galaxy of hundreds of high-velocity clouds with low line-of-sight velocities and small W50 widths also provokes the inclusion of false “nearby” dwarf galaxies in the LV"(@karachentsevUPDATEDNEARBYGALAXY2013)

@@

2.  Astro-Spam and Survey Errors:
    -   Automatic surveys produce false detections (e.g., stars misclassified as galaxies, high-velocity clouds mistaken for dwarfs).

3.  Conditional Completeness Estimate: - Galaxies brighter than $M_B^c = -11^m$ or with linear diameters larger than $A_{26} = 1.0$ kpc show (40–60)% completeness.
    -   "among the members of the Local Group (D \< 1 Mpc), only half of the galaxies have absolute magnitudes brighter than $−11^m$. Consequently, more than half of the ultra-faint dwarf companions around normal galaxies, like the Sombrero galaxy (D = 9.3 Mpc), still remain outside our field of view."

4.  Undetected Ultra-Faint Dwarfs:
    -   Many faint dwarf galaxies remain undetected beyond the Local Group. Estimated population of undetected dwarfs could be as large as $10^3–10^4$ within the LV.

5.  Surface Brightness Distribution:
    -   Surface brightness remains consistent across distances, except for ultra-faint dwarfs ($SB <31 \text{mag}\cdot\text{arcsec}^{-2}$).
    -   Faintest dwarf galaxies are detectable only nearby due to their resolution into individual stars.

6.  Luminosity-Size Relationship:
    -   Observations align with cosmological models predicting $L\sim A^3$ (@navarroStructureColdDark1996), though deviations occur for extremely low surface brightness galaxies.
    -   "The deviation from it at the extremely low surface brightness end is due to a systematic overestimation of dwarf galaxy sizes, the brightness profiles of which lie entirely below the Holmberg isophote."

@@

```{python}
#| output: asis
# Creating a dataframe with the lengths of each table
data = {
    "Table": ["Inner join", "Outer join", "LVG", "HECATE", "Unique galaxies in LVG", "Unique Galaxies in Hecate"],
    "Number of galaxies": [len(dt), len(outer), len(lvg), len(hec), len(lvg_not_hec), len(hec_not_lvg)]
}

@@

Completeness (LVG)= `{python} round(len(dt)/len(lvg) *100)` %

@@

Completeness (HECATE)= `{python} round(len(hec_not_lvg)/len(hec)*100)` %

Completeness (LVG)= `{python} round(len(lvg_not_hec)/len(lvg)*100)` %

Combined Completeness =$\frac{\text{Total galaxies in Outer}}{\text{Unique galaxies in HECATE + LVG}}$= `{python} round(len(outer)/(len(lvg_not_hec)+(len(hec_not_lvg))) * 100)` %

@@

```{python}
def table_completeness(dataset_choice, column_name="Dis", description = "Distance", keep_nan = False):
    """
    Creates a table of counts and percentages for each bin based on the chosen dataset (LVG or HEC).

    Parameters:
    - dataset_choice: A string indicating which dataset to use ("lvg" or "hec")
    - column_name: The column name in the datasets to analyze (default is "Dis")
    - description: what to show in the title and the axis (default "Distance")
    """

    if dataset_choice == "lvg":
        main_df, unique_df, label = lvg, lvg_not_hec, "LVG"
    elif dataset_choice == "hec" and hec is not None and hec_not_lvg is not None:
        main_df, unique_df, label = hec, hec_not_lvg, "HEC"
    else:
        raise ValueError("Invalid dataset choice. Please choose 'lvg' or 'hec' with valid datasets.")

@@

    # Calculate percentages in a vectorized way
    with np.errstate(divide='ignore', invalid='ignore'):  # Handle division by zero safely
        unique_pct = np.divide(unique_counts, main_counts, out=np.zeros_like(unique_counts, dtype=float), where=main_counts!=0) * 100
        inner_pct = np.divide(inner_counts, main_counts, out=np.zeros_like(inner_counts, dtype=float), where=main_counts!=0) * 100

    # Construct the DataFrame in a vectorized manner
    table_df = pd.DataFrame({
        "Bin Start": np.round(bin_edges[:-1]),
        "Bin End": np.round(bin_edges[1:]),
        "Main Counts": main_counts,
        "Unique Counts": unique_counts,
        "Unique %": unique_pct,
        "Inner Counts": inner_counts,
        "Inner %": inner_pct
    })
    return table_df
```

```{python}
#| fig-cap: "Histograms showing the Distance Completeness of the Catalogs"
#| layout-ncol: 2
#| fig-subcap: 
#| - "HECATE"
#| - "LVG"
#| label: fig-dis-comp

@@

main = hec["D"].data
unique = hec_not_lvg["D"].data
counts, bin_edges = np.histogram(main)

@@

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)
ax2.set(xlabel = "Distance [{}]".format(lvg["Dis"].unit), ylabel="Completeness %")
plt.show()

#LVG
main = lvg["Dis"].data
unique = lvg_not_hec["Dis"].data
counts, bin_edges = np.histogram(main)

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "LVG", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LVG", ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Distance for LVG")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100 
###
sns.lineplot(x = bin_edges[1:], y = unique_pct, marker = "o", ax = ax2)

@@

# Format the y-tick labels to display as percentages
y_value = ['{:,.0f}%'.format(x) for x in ax2.get_yticks()]
ax2.set_yticklabels(y_value)
ax2.set(xlabel = "Distance [{}]".format(lvg["Dis"].unit), ylabel="Completeness %")
plt.show()

@@

```{python}
#| fig-cap: "Histograms showing the Type Completeness of the Catalogs"
#| layout-ncol: 2
#| fig-subcap: 
#| - "HECATE"
#| - "LVG"
#| label: fig-type-comp
main = hec["T"].data
main = main[~np.isnan(main)]
unique = hec_not_lvg["T"].data
unique = unique[~np.isnan(unique)]
counts, bin_edges = np.histogram(main, bins = 12)

@@

ax2.set(xlabel = "Types of galaxies", ylabel="Completeness %")
plt.show()
######
#LVG
main = lvg["TType"].data
main = main[~np.isnan(main)]
unique = lvg_not_hec["TType"].data
unique = unique[~np.isnan(unique)]
counts, bin_edges = np.histogram(main)

@@

fig, (ax1, ax2) = plt.subplots(2, 1, sharex=True)
sns.histplot(main, bins = bin_edges, kde = True, label = "LVG", ax = ax1)
sns.histplot(unique, bins = bin_edges, kde = True, label = "Unique for LVG", ax = ax1)
ax1.legend()
ax1.set_title("Completenes of Types for LVG")
ax1.tick_params(labelbottom=True)
unique_pct = (unique_counts/main_counts)*100

@@

```{=html}
<!--As we can see from the histograms (@fig-dis-comp) and (@fig-type-comp) the sample of unique galaxies of each catalog, gets smaller by an almost constant proportion (Inner join).

This means there is no bias in the selection of the galaxies.-->
```
# How are we going to compare the data?

@@

For the distribution comparisons we will use histograms and Kernel Density Estimate (KDE) plots. The KDE plots visually represent the distribution of data, providing insights into its shape, central tendency, and spread.

@@

$$\text{Percentage change} = \frac{V_{Hecate} - V_{LVG}}{V_{Hecate}}\cdot 100 \%$$

```{python}
def relative_diff(hec, lvg):
    """
    Calculate the relative difference between two sets of values.

    Parameters:
    hec (numpy.ndarray): An array of values representing the HEC dataset.
    lvg (numpy.ndarray): An array of values representing the LVG dataset.

    Returns:
    numpy.ndarray: An array of relative differences between the HEC and LVG datasets.
    The relative difference is calculated as ((HEC - LVG) * 100) / |HEC|.
    Any infinite values in the result are replaced with NaN.
    """
    diff = (hec - lvg) * 100 / np.abs(hec)
    diff[np.isinf(diff)] = np.nan
    return diff
```

@@

| LVG | HECATE | Description |
|:---------------:|:---------------:|:---------------:|
| Dis |   D    |  Distance   |

@@

plt.close()
fig, ax = plt.subplots()
sns.regplot(x = dt["Dis"].data, y = dt["D"].data,
                 scatter_kws=dict(alpha=0.5, color='blue', edgecolors='white'),
                 line_kws=dict(alpha=0.7, color='red', linewidth=3))
ax.errorbar(dt["Dis"].data, dt["D"].data, yerr = dt["E_D"].data, fmt='none', capsize=2, zorder=1, color = "blue", alpha = 0.3)
results = scipy.stats.linregress(x = dt["Dis"], y=dt["D"])
plt.xlabel("$D_{LVG}$"+" [{}]".format(dt["Dis"].unit))
plt.ylabel("$D_{HEC}$"+" [{}]".format(dt["D"].unit))
plt.legend(["Data",
            "$D_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot D_{LVG}+$"+
                f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.tight_layout()
plt.show()
```

@@

|     LVG     |    HECATE    |              Description               |
|:----------------:|:-----------------:|:----------------:|
| RVel (km/s) |   V (km/s)   |      Heliocentric radial velocity      |
| VLG (km/s)  |              |            Radial velocity             |
|  cz (km/s)  |              |         Heliocentric velocity          |
|             | V_VIR (km/s) | Virgo-infall corrected radial velocity |

@@

ax.errorbar(dt["RVel"].data, dt["V"].data, yerr = dt["E_V"].data, fmt='none', capsize=2, zorder=1, color = "blue")
results = scipy.stats.linregress(x = dt["RVel"], y=dt["V"])
plt.xlabel("$V_{LVG}$"+" [{}]".format(dt["RVel"].unit))
plt.ylabel("$V_{HEC}$"+" [{}]".format(dt["V"].unit))
plt.legend(["Data",
            "$V_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot V_{LVG}+$"+
            f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()
```

@@

```{python}
#| label: fig-vel-pairplot
#| fig-cap: "The correlation Matrix of the Velocities. The lower left triangle is composed of the scatter plots of the various velocities, the diagonal shows their distrubution and the upper right triangle shows the correlations of the Velocities" 
pairplot_with_correlation(vel_data)
```

\[?\] The strong correlation between the velocities reported by the HECATE and LVG catalogs (@fig-vel-pairplot) can be attributed to the fact that both measure the same intrinsic galaxy velocities but reference them to different frames. For instance, velocities can be corrected for the movement of the Earth (heliocentric frame) or the for the Virgo-infall. These corrections shift the measured velocity slightly, but the fundamental measurement remains consistent. This explains why the velocities are highly correlated, as they represent the same physical phenomenon but adjusted for different reference points.

{{< pagebreak >}}

@@

|      LVG      |        HECATE        |                        Description                        |
|:----------------:|:----------------:|:-----------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |
|      inc      |         INCL         |                     Inclination (deg)                     |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |

@@

"as they correspond to types T \< 1. The shortcomings of such a simplified classification of dwarf systems have become apparent; hence, van den Bergh (1959) proposed a more refined scheme where dwarf galaxies were assumed to be divided by luminosity class...Such an approach allowed us to avoid unpleasant cases in which a dwarf galaxy of intermediate type jumps over due to a misclassification from one end of the de Vaucouleurs sequence (T \< 1) to the other (T = 9, 10)." (@karachentsevSTARFORMATIONPROPERTIES2013a)

```{python}
#| label: fig-types-compare
#| layout-ncol: 2
#| layout-align: center
#| fig-cap: "Comparison of the Types of galaxies"
#| fig-subcap: 
#| - "All the galaxies, without the correction" 
#| - "Comparison with the correction $|{T_{HECATE}-T_{LVG}}|<σ_{Τ_{LVG}}$"

@@

ax.errorbar(dt["TType"].data, dt["T"].data, yerr = dt["E_T"].data, fmt='none', capsize=2, zorder=1, color = "blue")
temp = dt[["TType", "T"]].to_pandas().dropna()
results = scipy.stats.linregress(x = temp["TType"],
                                y = temp["T"])
plt.xlabel("$Type_{LVG}$")
plt.ylabel("$Type_{HEC}$")
plt.legend(["Data",
            "$T_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot T_{LVG}+$"+
            f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()

@@

ax.errorbar(types["TType"].data, types["T"].data, yerr = types["E_T"].data, fmt='none', capsize=2, zorder=1, color = "blue")
results = scipy.stats.linregress(x= types["TType"].data,
                                y = types["T"].data)
plt.xlabel("$Type_{LVG}$")
plt.ylabel("$Type_{HEC}$")
plt.legend(["Data",
            "$T_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot T_{LVG}+$"+
            f"$({results.intercept:+.2f})$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()
```

-   we can assume that the galaxies from the upper left and lower right regions of the plot are classified differently because of this
-   this might explain the large errors $\delta T$ of HECATE
-   why we see a difference in the distribution of (@fig-type-comp)

We can remove the "problematic" galaxies by only keeping the ones with: 
$$
|{T_{HECATE}-T_{LVG}}|<\sigma_{T_{LVG}} = 4.8
$$

-   The average uncertainty of the morphological type in the HECATE catalog is $\overline{\delta T} = \pm 1.4$, so the intercept is included in the error.
-   So we can assume that the Morphological Types are the same

@@

sns.histplot(temp["INCL"], kde = True, label = "HECATE")
sns.histplot(temp["inc"], kde = True, label = "LVG")
plt.legend()
plt.xlabel("Inclination of galaxies [deg]")
plt.show()

@@

```{python}
# Display specific statistics, excluding the 25th and 75th percentiles
summary_stats = temp.describe().drop(['25%', '75%'])
summary_stats
```

We can see that for values in the range $[\sim 30^\circ,\sim 80^\circ]$, the values of the LVG inclination are higher (@fig-incl-compare). However, since their means, median, min and maxes are similar and the percentage change is practically 0% (mean, median, $\sigma$ = 0 with a range $[-3\%,1\%]$), we can ignore the differences and assume they are the same values.

@@

plt.xlabel("(Major angular diameter$)_{LVG}$")
plt.ylabel("(Semi major axis$)_{HEC}$")
plt.legend(["Data",
            "$R1_{HEC}=$" +f"{results.slope:.2f}"+"$\cdot a_{26,LVG}$"+
            f"${results.intercept:+.2f}$"+
            f"\n $R^2 = {results.rvalue**2*100:.2f}\%$"])
plt.show()
dt["log_a26"] = np.log10(dt["a26_1"].data)
dt["log_R1"] = np.log10(dt["R1"].data)
plt.close()
fig, ax = plt.subplots()
p = sns.regplot(x = dt["log_a26"].data, y = dt["log_R1"].data,
                 scatter_kws=dict(alpha=0.5, color='blue', edgecolors='white'),
                 line_kws=dict(alpha=0.7, color='red', linewidth=3))

temp = dt[["log_a26","log_R1"]].to_pandas().dropna()
results = scipy.stats.linregress(x = temp["log_a26"], y = temp["log_R1"])
plt.xlabel("(Major angular diameter$)_{LVG}$"+" [{}]".format(dt["a26_1"].unit))
plt.ylabel("(Semi major axis$)_{HEC}$"+" [{}]".format(dt["R1"].unit))
plt.legend(["Data",
            "$log_{10}(R1_{HEC})=$" +f"{results.slope:.2f}"+"$\cdot log_{10}(a_{26,LVG})$"+
            f"${results.intercept:+.2f}$"+
            f"\n $R^2 = {results.rvalue**2*100:.0f}\%$"])
plt.show()
```
